### 2016-07-15 Introduction.tex
COMMITTED VERSION

\section{Introduction}
The Standard Model of particle physics is a quantum field theory which relies on the principle of gauge invariance. It describes the world's fundamental particles and the interactions among them and predicts the most accurately measured observables in physics.\\
Besides these triumphs, it is known to be incomplete for example for not including dark matter or a description of gravity. Because of these and other flaws, one studies extensions of the Standard Model. One mathematically very appealing extension is the concept of supersymmetry. Within this framework one often studies the Minimal Supersymmetric Standard Model (MSSM). This theory provides a candidate for dark matter and elegant solutions of e.g. the hierarchy problem. However the MSSM does not come without problems. One of these is that supersymmetric particles have not been discovered so far. This is  actually reasoned by introducing supersymmetry breaking. However, in order to still solve the hierarchy problem, supersymmetry breaking should not be too large. According to searches for supersymmetry at the Large Hadron Collider, supersymmetric particles are at the moment on the verge on being too heavy for the MSSM to solve the hierarchy problem without loosing its attractiveness.\\
A non-trivial extension of supersymmetry - $R$-symmetry - may explain this and other issues. $R$-symmetry is, according to the Haag-Lopuszanski-Sohnius theorem, the only viable extension of the supersymmetry group of the MSSM. This is, the combination of Poincaré symmetry, supersymmetry and $R$-symmetry is the maximal symmetry a reasonable quantum field theory can have.\\
\ignore{Imposing supersymmetry and $R$-symmetry on the Standard Model, gives - according to the Haag-Lopuszanski-Sohnius theorem - the maximal symmetry group a reasonable quantum field theory can have.\\}
Motivated be these facts, the Minimal $R$-Symmetric Supersymmetric Standard Model (MRSSM) is the subject of investigations within this thesis. More specifically, the production cross section of strongly interacting supersymmetric particles in the MRSSM is studied. It will turn out that there are significant changes in the production of squarks and gluinos when turning from the MSSM to the MRSSM.\\
This thesis starts with summarizing the main aspects of the Standard Model before the concept of supersymmetry and $R$-symmetry is introduced. This includes the introduction of the models which incorporate the minimal form of the symmetry in question. Section \ref{sec:SquarkGluinoTree} compares the production cross section of squark and gluinos in both models at \mbox{tree-level}. After identifying an attractive region of parameter space in the MRSSM, it will turn out that the most interesting difference manifests in the channel of squark production. As there are no programs available with whom next-to-leading order cross sections within the MRSSM can be calculated, the rest of the thesis focuses on the next-to-leading order calculation of the squark production cross section.
: Section \ref{sec:VirtRealCorr} discusses the golden thread of the calculation of a next-to-leading order cross section including massless external particles before section \ref{sec:renMRSSM} gives detailed information about the renormalization of the MRSSM. This includes the explicit form of required renormalization constants, reaching form the on-shell scheme over a mixture of $\overline{\mathrm{MS}}$- and zero-momentum subtraction scheme to supersymmetry restoring counterterms. This section also discusses how renormalized one-loop matrix elements are generated as an input for a program which calculates the squark production cross section in the MRSSM.
The thesis closes with an explanation on how this very program is constructed and the representation of the next-to-leading order cross section for squark production in the form of $K$-factors.

>>>>

\section{Introduction}
The Standard Model of particle physics is a quantum field theory which relies on the principle of gauge invariance. It describes all known fundamental particles and the non-gravitational interactions among them and predicts the most accurately measured observables in physics.\\
Besides these successes, it is known to be incomplete for example for not including dark matter or a description of gravity. Because of these and other flaws, one studies extensions of the Standard Model. One mathematically very appealing extension is the concept of supersymmetry. Within this framework one often studies the Minimal Supersymmetric Standard Model (MSSM). This theory may provide a candidate for dark matter and elegant solutions of e.g. the hierarchy problem. However, the MSSM does not come without problems. One of these is that supersymmetric particles have not been discovered so far. This is  actually reasoned by introducing supersymmetry breaking. However, in order to still solve the hierarchy problem, supersymmetry breaking should not be too large. According to searches for supersymmetry at the Large Hadron Collider (LHC), supersymmetric particles are at the moment on the verge of being too heavy for the MSSM to solve the hierarchy problem without losing its attractiveness.\\
A non-trivial extension of supersymmetry -- $R$-symmetry -- may explain this and other issues. $R$-symmetry is, according to the Haag-{\L}opusza\'nski-Sohnius theorem, the only conceivable extension of the symmetry group of the MSSM. This is, the combination of Poincaré symmetry, supersymmetry and $R$-symmetry is the maximal symmetry a reasonable quantum field theory can have.\\
\ignore{Imposing supersymmetry and $R$-symmetry on the Standard Model, gives - according to the Haag-{\L}opusza\'nski-Sohnius theorem - the maximal symmetry group a reasonable quantum field theory can have.\\}
Motivated by these facts, the Minimal $R$-Symmetric Supersymmetric Standard Model (MRSSM) is the subject of investigations within this thesis. More specifically, the production cross section of strongly interacting supersymmetric particles in the MRSSM is studied. This is because within supersymmetry strongly interacting particles are the ones which are most likely to be produced at the LHC should supersymmetry at all be realized in nature. It will turn out that there are significant changes in the production of squarks and gluinos when turning from the MSSM to the MRSSM.\\
This thesis starts with summarizing the main aspects of the Standard Model before the concepts of supersymmetry and $R$-symmetry are introduced. This includes the introduction of the models which incorporate the minimal form of the symmetry in question. Section \ref{sec:SquarkGluinoTree} compares the production cross sections of squarks and gluinos in both models at \mbox{tree-level}. After identifying an attractive region of parameter space in the MRSSM, it will turn out that the most interesting difference manifests in the channel of squark production. As there are no computer programs available with whom next-to-leading order cross sections within the MRSSM can be calculated, the rest of the thesis focuses on the next-to-leading order calculation of the squark production cross section. Section \ref{sec:VirtRealCorr} outlines the structure of the calculation of a next-to-leading order cross section including massless external particles before section \ref{sec:renMRSSM} gives detailed information about the renormalization of the MRSSM. This includes the explicit form of required renormalization constants, reaching from the on-shell scheme over a mixture of $\overline{\mathrm{MS}}$- and zero-momentum subtraction scheme to supersymmetry restoring counterterms. This section also discusses how renormalized one-loop matrix elements are generated as an input for a program which calculates the squark production cross section in the MRSSM.
The thesis closes with an explanation of how this very program is constructed and the representation of the next-to-leading order cross section for squark production in the form of $K$-factors.

### 2016-07-15 MSSM.tex
COMMITTED VERSION

\subsection{Supersymmetry as an Extention of the Poincaré Symmetry}\label{sec:SUSYalgebra}
As already seen in the previous chapter, symmetries are of vital importance when studying particle physics. Coleman and Mandula \cite{Coleman:1967ad} found that for any reasonable quantum field theory space-time and internal symmetries, i.e. gauge symmetries ``cannot by combined in any but the trivial way''\cite{Pelc:1996vg}. Ergo, the symmetry groups of the Standard Model are the most general ones.\\
This statement was corrected by the Haag-Lopuszanski-Sohnius theorem \cite{Haag:1974qh} which weakened the condition of the quantum field theory by allowing also for anticommuting symmetry generators. This is, by introducing the generators $Q_\alpha$ and $\overline{Q}^{\dot{\alpha}}$ with $\alpha, \dot{\alpha} \in \left\{ 1,2 \right\}$, the Poincare-algebra eq. \eqref{eq:PoincareAlgebra} can be nontrivially extended to the so called superalgebra (see Appendix \eqref{sec:2spinor_notation} for the definition of the $\sigma$-matrices):
\begin{align}
\left\{Q_\alpha,Q_\beta\right\} = \left\{\overline{Q}_{\dot{\alpha}},\overline{Q}_{\dot{\beta}}\right\} &= 0,  \nonumber\\
\left\{Q_\alpha,\overline{Q}_{\dot{\alpha}}\right\} &= 2\sigma^\mu_{\alpha\dot{\alpha}} P_\mu, \nonumber\\
[P^\mu,Q_\alpha] = [P^\mu,\overline{Q}_{\dot{\alpha}}] &= 0, \nonumber \\
[Q_\alpha, J^{\mu\nu}] &= \frac{1}{2} (\sigma^{\mu\nu})_\alpha^{\ \beta} Q_\beta.\label{eq:SUSYalgebra}
\end{align}
As $J^{\mu\nu}$ is the generalized angular momentum operator, the last line in eq. \eqref{eq:SUSYalgebra} implies that $Q_\alpha$ is a spin-$\frac{1}{2}$ operator. Applying $Q_\alpha$ to a field will change the spin of it by $\frac{1}{2}$. This means supersymmetry generators convert fermions to bosons and vice versa.\\
The superalgebra in eq. \eqref{eq:SUSYalgebra} is actually the smallest extension of the Poincare-algebra as it is possible to introduce further anticommuting symmetry generators. Their maximal number is fixed to $N = 8$. Because the MSSM is a minimal extension of the Standard Model, it deals with only one set of anticommuting symmetry generators and is therefore referred to be a $N = 1$ supersymmetric theory.\\
A representation in the form of differential operators can be given by introducing superspace. Superspace is a manifold obtained by enlarging Minkowski space, whose coordinates are labeled with $x^\mu$, with four anticommuting numbers: $\theta^\alpha$ and $\overline{\theta}^{\dot{\alpha}}$, where $\alpha,\ \dot{\alpha} \in \left\{ 1,2 \right\}$ (see Appendix \ref{sec:AnticommNumbers} for the definition of anticommuting numbers). A representation of the superalgebra reads
\begin{align}
P^\mu &= i\partial^\mu,\nonumber\\
J^{\mu\nu} &= i(x^\mu\partial^\nu - x^\nu\partial^\mu),\nonumber\\
Q_\alpha &= i(\partial_\alpha + i\sigma^\mu_{\alpha\dot{\alpha}}
\overline{\theta}^{\dot{\alpha}}\partial_\mu),\nonumber\\
\overline{Q}_{\dot{\alpha}} &= i(-\overline{\partial}_{\dot{\alpha}} - i \theta^\alpha \sigma^\mu_{\alpha\dot{\alpha}}\partial_\mu).\label{eq:SUSYGen}
\end{align}

@@

Note that the ``chirality'' of sfermions is understood to be the chirality of their superpartners. Of course scalar particles have no handedness.
\begin{table}

>>>>

\subsection{Supersymmetry as an Extention of the Poincaré Symmetry}\label{sec:SUSYalgebra}
As already seen in the previous chapter, symmetries are of vital importance when studying particle physics. Coleman and Mandula \cite{Coleman:1967ad} found that for any reasonable quantum field theory space-time and internal symmetries, i.e. gauge symmetries ``cannot by combined in any but the trivial way''\cite{Pelc:1996vg}. Ergo, the symmetry groups of the Standard Model are the most general ones.\\
This statement was corrected by the Haag-{\L}opusza\'nski-Sohnius theorem \cite{Haag:1974qh} which weakened the condition of the quantum field theory by allowing also for anticommuting symmetry generators. This is, by introducing the generators $Q_\alpha$ and $\overline{Q}^{\dot{\alpha}}$ with $\alpha, \dot{\alpha} \in \left\{ 1,2 \right\}$, the Poincare-algebra eq. \eqref{eq:PoincareAlgebra} can be nontrivially extended to the so called superalgebra (see Appendix \eqref{sec:2spinor_notation} for the definition of the $\sigma$-matrices):
\begin{align}
\left\{Q_\alpha,Q_\beta\right\} = \left\{\overline{Q}_{\dot{\alpha}},\overline{Q}_{\dot{\beta}}\right\} &= 0,  \nonumber\\
\left\{Q_\alpha,\overline{Q}_{\dot{\alpha}}\right\} &= 2\sigma^\mu_{\alpha\dot{\alpha}} P_\mu, \nonumber\\
[P^\mu,Q_\alpha] = [P^\mu,\overline{Q}_{\dot{\alpha}}] &= 0, \nonumber \\
[Q_\alpha, J^{\mu\nu}] &= \frac{1}{2} (\sigma^{\mu\nu})_\alpha^{\ \beta} Q_\beta.\label{eq:SUSYalgebra}
\end{align}
As $J^{\mu\nu}$ is the generalized angular momentum operator, the last line in eq. \eqref{eq:SUSYalgebra} implies that $Q_\alpha$ is a spin-$\frac{1}{2}$ operator. Applying $Q_\alpha$ to a field will change the spin of it by $\frac{1}{2}$. This means supersymmetry generators convert fermions to bosons and vice versa.\\
The superalgebra in eq. \eqref{eq:SUSYalgebra} is actually the smallest extension of the Poincare-algebra as it is possible to introduce further anticommuting symmetry generators. Their maximal number is fixed to $N = 8$. Because the MSSM is a minimal extension of the Standard Model, it deals with only one set of anticommuting symmetry generators and is therefore referred to be a $N = 1$ supersymmetric theory.\\
A representation in the form of differential operators can be given by introducing superspace. Superspace is a manifold obtained by enlarging Minkowski space, whose coordinates are labeled with $x^\mu$, with four anticommuting numbers: $\theta^\alpha$ and $\overline{\theta}^{\dot{\alpha}}$, where $\alpha,\ \dot{\alpha} \in \left\{ 1,2 \right\}$ (see Appendix \ref{sec:AnticommNumbers} for the definition of anticommuting numbers). A representation of the superalgebra reads
\begin{align}
P^\mu &= i\partial^\mu,\nonumber\\
J^{\mu\nu} &= i(x^\mu\partial^\nu - x^\nu\partial^\mu),\nonumber\\
Q_\alpha &= i(\partial_\alpha + i\sigma^\mu_{\alpha\dot{\alpha}}
\overline{\theta}^{\dot{\alpha}}\partial_\mu),\nonumber\\
\overline{Q}_{\dot{\alpha}} &= i(-\overline{\partial}_{\dot{\alpha}} - i \theta^\alpha \sigma^\mu_{\alpha\dot{\alpha}}\partial_\mu).\label{eq:SUSYGen}
\end{align}

@@

Note that the ``chirality'' of sfermions is understood to be the chirality of their superpartners. Of course scalar particles have no chirality.
\begin{table}[H]

### 2016-07-15 Masterarbeit.tex
Summary chapter
K factor MSSM contour plot
COMMITTED VERSION

\pagenumbering{gobble}   % no pagenumbering
% Titelpageseite
\begin{titlepage}
 \begin{tabularx}{\linewidth}{X}
  \includegraphics[width=6cm]{TU_Logo_SW} \\\hline\hline
  \vspace{4.5em}
  \begin{singlespace}\begin{center}\bfseries\Huge  
  Squark Production in the R-Symmetric Supersymmetric Standardmodel  
  \end{center}\end{singlespace}
  \vspace{5.5em}
  \begin{singlespace}\begin{center}\large
   Master-Arbeit \\ zur Erlangung des Hochschulgrades \\ 
   Master of Science \\ 
   im Master-Studiengang Physik
  \end{center}\end{singlespace}\medskip
  \begin{center}vorgelegt von\end{center}
  \begin{center}
   {\large Sebastian Liebschner} \\ geboren am 08.05.1990 in Meerane
  \end{center}\medskip
  \begin{singlespace}\begin{center}\large
   Institut für Kern- und Teilchenphysik \\
   Fachrichtung Physik \\
   Fakultät Mathematik und Naturwissenschaften \\
   Technische Universität Dresden \\ 2016
  \end{center}\end{singlespace}
 \end{tabularx}
\end{titlepage}

@@

% abstract page
\begin{center}\large\bfseries Summary\end{center}
Abstract \\ 
English: flaws of the SM\\
motivation: aesthetic: Coleman-Mandula --> Haag-Lopuszanski-Sohnius-Theorem\\
\vspace{20em}
Abstract \\ 
Deutsch \\
\newpage

@@

\pagenumbering{arabic}
\input{Introduction}
\newpage
\input{Standardmodel}
\newpage
\input{MSSM}
\newpage
\input{RSymmetry}
\newpage
\input{TreeLevel}
\newpage
\input{Virt_Real}
\newpage
\input{Renormalization}
\newpage
\input{NLO}
\newpage
\section{Summary and Outlook}
\newpage
\input{Appendix}
\newpage
\bibliography{references}
\newpage
% Erklärung
\clearpage
\thispagestyle{empty}
\minisec{Erklärung}\vspace*{1.5em}
Hiermit erkläre ich, dass ich diese Arbeit im Rahmen der Betreuung am Institut
für Kern- und Teilchenphysik ohne unzulässige Hilfe Dritter verfasst und alle Quellen als solche gekennzeichnet habe.
\vspace*{45em}
Vorname Nachname \par
Dresden, Juli 2016

>>>>

\pagenumbering{gobble}   % no pagenumbering
% Titelpageseite
\begin{titlepage}
 \begin{tabularx}{\linewidth}{X}
  \includegraphics[width=6cm]{TU_Logo_SW} \\\hline\hline
  \vspace{4.5em}
  \begin{singlespace}\begin{center}\bfseries\Huge  
  Squark Production in the R-Symmetric Supersymmetric Standardmodel  
  \end{center}\end{singlespace}
  \vspace{5.5em}
  \begin{singlespace}\begin{center}\large
   Masterarbeit \\ zur Erlangung des Hochschulgrades \\ 
   Master of Science \\ 
   im Master-Studiengang Physik
  \end{center}\end{singlespace}\medskip
  \begin{center}vorgelegt von\end{center}
  \begin{center}
   {\large Sebastian Liebschner} \\ geboren am 08.05.1990 in Meerane
  \end{center}\medskip
  \begin{singlespace}\begin{center}\large
   Institut für Kern- und Teilchenphysik \\
   Fachrichtung Physik \\
   Fakultät Mathematik und Naturwissenschaften \\
   Technische Universität Dresden \\ 2016
  \end{center}\end{singlespace}
 \end{tabularx}
\end{titlepage}

@@

% abstract page
%\begin{center}\large\bfseries Summary\end{center}
\section*{\centering{Kurzdarstellung}}
Im Streben der Physik, die fundamentalen Bausteine der Natur zu beschreiben, stellt die Ent\-wicklung des Standardmodells der Teilchenphysik einen Meilenstein dar. Dieses beschreibt alle bekannten Elementarteilchen sowie deren nicht-gravitative Wechselwirkungen. Neben diesen Erfolgen ist aber auch bekannt, dass es unvollständig ist, da es keine Gravitation oder dunkle Materie beinhaltet. Supersymmetrie ist eine attraktive Erweiterung des Standardmodells, weil sie viele Probelme desselben auf eine elegante Weise löst.\\
Diese Arbeit untersucht eine supersymmetrische Erweiterung des Standardmodells welche zusätzlich $R$-symmetrisch ist. Neben der Berechnung von Wirkungsquerschnitten von stark wechelwirkenden supersymmetrischen Teilchen in der führenden Ordnung werden Einschleifen\-korrekturen zum Produktionswirkungsquerschnitt von Squarks in diesem Model berechnet und phenomenologisch ausgewertet.
\vspace{10em}
\section*{\centering{Abstract}}
In the ambition of understanding the fundamental mechanisms of nature, the Standard Model of particle physics constitutes a cornerstone in physics. It describes all known elementary particles and their non-gravitational interactions. Besides these successes it is known to be incomplete since it does not include gravity or dark matter. Supersymmetry is an attractive extension of the Standard Model as it solves many problems of it elegantly.\\
Within this thesis, a supersymmetric model which also exhibits $R$-symmetry is considered. Besides the calculation of tree-level cross sections for the production of strongly interacting supersymmetric particles, also one-loop corrections to the production cross section of squarks are calculated and phenomenologically analyzed.
\newpage

@@

\pagenumbering{arabic}
\input{Introduction}
\newpage
\input{Standardmodel}
\newpage
\input{MSSM}
\newpage
\input{RSymmetry}
\newpage
\input{TreeLevel}
\newpage
\input{Virt_Real}
\newpage
\input{Renormalization}
\newpage
\input{NLO}
\newpage
\input{Summary}
\newpage
% Leerseite
%\thispagestyle{empty}
\cleardoublepage

%\appendix
\input{Appendix}
\newpage
\bibliography{references}
\newpage
\section*{\centering{Acknowledgements}}
First of all I like to thank Dominik Stöckinger who enthused me with theoretical particle physics and provided me with this interesting topic. In addition, he was always available for helping discussions whenever I was in need of them.\\
Special thanks go to Philip Dießner and Wojciech Kotlarski, who contributed significantly to results within this thesis. In addition I like to thank them, Markus Bach, Thomas Kwasnitza, Josua Unger and Tom Steudtner for many valuable discussions and support regarding physical questions as well as programming issues and a pleasant atmosphere in the office. Furthermore, I like to give them thanks for proofreading this thesis.\\
I am particularly grateful for Rebecca Schmidt's support within the last month which enabled me to focus on my thesis. In addition I like to thank her for proofreading.\\
Last but not least I would like to thank my family. Without them my physics studies would by no means have been possible in the way they were.
\newpage
% Erklärung
\clearpage
\thispagestyle{empty}
\minisec{Erklärung}\vspace*{1.5em}
Hiermit erkläre ich, dass ich diese Arbeit im Rahmen der Betreuung am Institut
für Kern- und Teilchenphysik ohne unzulässige Hilfe Dritter verfasst und alle Quellen als solche gekennzeichnet habe.
\vspace*{45em}
\par
Sebastian Liebschner\\
Dresden, Juli 2016

### 2016-07-15 RSymmetry.tex
COMMITTED VERSION

\section{$R$-Symmetry}
$R$-symmetry is an additional continuous symmetry which extends  supersymmetry in a non-trivial way. This is actually covered in the Haag-Lopuszanski-Sohnius theorem \cite{Haag:1974qh} but often concealed.\\
This section introduces $R$-symmetry and discusses briefly a minimal viable extension of the MSSM including $R$-symmetry, the MRSSM. Thereby some general features of the model are mentioned before the focus is laid on the strongly interacting sector which is the central topic of the thesis.

>>>>

\section{$R$-Symmetry}
$R$-symmetry is an additional continuous symmetry which extends  supersymmetry in a non-trivial way. This is actually covered in the Haag-{\L}opusza\'nski-Sohnius theorem \cite{Haag:1974qh} but often concealed.\\
This section introduces $R$-symmetry and discusses briefly a minimal viable extension of the MSSM including $R$-symmetry, the MRSSM. Thereby some general features of the model are mentioned before the focus is laid on the strongly interacting sector which is the central topic of the thesis.
